---
title: "FDS Final Project: Report #2"
output: html_document
---

# Final Project # 2: ZURICH POLITICIANS DATA ANALYSIS

## Analysis of Politicians data from Zurich from a SQLite database using R

### Loading required Libraries 

## Connecting Database and Cleaning


```{r}

library(DBI)
library(RSQLite)
library(dplyr)
library(tibble)
library(tidyverse)

# Connect to database and suppress type coercion warnings from SQLite
conn <- dbConnect(SQLite(), "zh_politicians.db")

suppressWarnings({
  persons <- dbGetQuery(conn, "SELECT * FROM PERSONS") %>% tibble()
  mandates <- dbGetQuery(conn, "SELECT * FROM MANDATES") %>% tibble()
  affiliations <- dbGetQuery(conn, "SELECT * FROM AFFILIATIONS") %>% tibble()
  addresses <- dbGetQuery(conn, "SELECT * FROM ADDRESSES") %>% tibble()
})

dbDisconnect(conn)

# 2. Clean and coerce numeric columns across tables in a single block
persons <- persons %>%
  mutate(
    YEAR_OF_BIRTH = as.numeric(YEAR_OF_BIRTH),
    YEAR_OF_DEATH = as.numeric(YEAR_OF_DEATH)
  )

mandates <- mandates %>%
  mutate(
    MANDATE_START_DAY = as.numeric(MANDATE_START_DAY),
    MANDATE_START_MONTH = as.numeric(MANDATE_START_MONTH),
    MANDATE_START_YEAR = as.numeric(MANDATE_START_YEAR),
    MANDATE_END_DAY = as.numeric(MANDATE_END_DAY),
    MANDATE_END_MONTH = as.numeric(MANDATE_END_MONTH),
    MANDATE_END_YEAR = as.numeric(MANDATE_END_YEAR)
  )

affiliations <- affiliations %>%
  mutate(
    AFFILIATION_START_DAY = as.numeric(AFFILIATION_START_DAY),
    AFFILIATION_START_MONTH = as.numeric(AFFILIATION_START_MONTH),
    AFFILIATION_START_YEAR = as.numeric(AFFILIATION_START_YEAR),
    AFFILIATION_END_DAY = as.numeric(AFFILIATION_END_DAY),
    AFFILIATION_END_MONTH = as.numeric(AFFILIATION_END_MONTH),
    AFFILIATION_END_YEAR = as.numeric(AFFILIATION_END_YEAR)
  )

cat("Step 1 executed cleanly without duplicates!\n")
```

# Part 1. Exploring the assembly composition


```{r}
library(purrr)
library(writexl)

# Filter valid mandate ranges and expand years into list-columns
mandates_expanded <- mandates %>%
  filter(
    !is.na(MANDATE_START_YEAR), 
    !is.na(MANDATE_END_YEAR), 
    MANDATE_START_YEAR <= MANDATE_END_YEAR
  ) %>%
  mutate(years = map2(MANDATE_START_YEAR, MANDATE_END_YEAR, seq)) %>%
  unnest(years)

# Calculate active mandates evolution per assembly over time
active_evolution <- mandates_expanded %>%
  group_by(ASSEMBLY, years) %>%
  summarise(active_count = n(), .groups = 'drop')

write_xlsx(active_evolution, "active_mandates_evolution.xlsx")

library(ggplot2)

# Generate a line plot of mandate trends across assemblies
ggplot(active_evolution, aes(x = years, y = active_count, color = ASSEMBLY)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Evolution of Active Mandates per Assembly",
    subtitle = "Tracking active mandates over the years (showing election-year bumps)",
    x = "Year",
    y = "Number of Active Mandates",
    color = "Assembly"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )
```

### REMARKS: Total dominance of Cantonal Council between the years of 1500 and 2000
### Cantonal Council has a massive expansion over a 400 active mandates



# Part 2. Gender Composition Evolution by Assembly


```{r}
# 3. Part 2: Calculate gender composition evolution per assembly over time
gender_evolution <- mandates_expanded %>%
  left_join(persons %>% select(ID, GENDER), by = c("PERSON_ID" = "ID")) %>%
  group_by(ASSEMBLY, years, GENDER) %>%
  summarise(count = n(), .groups = 'drop')

write_xlsx(gender_evolution, "gender_composition_evolution.xlsx")

library(ggplot2)
library(dplyr)

gender_evolution %>%
  ggplot(aes(x = years, y = count, color = GENDER)) +
  geom_line(linewidth = 1) +
  facet_wrap(~ASSEMBLY, scales = "free") + 
  labs(
    title = "Gender Composition Evolution by Assembly",
    subtitle = "Comparing active mandates held by women and men over time",
    x = "Year",
    y = "Number of Active Mandates",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )
```

### Gran Council shows a continuous line of male representation across the centuries

### The Cantonal Council and Executive Council reflects how female participation increase toward the end of 20th century and early 21st century


# Part 3. Party Composition in the year 2000 

```{r}
library(dplyr)
library(ggplot2)
library(writexl)

# 1. Filter active politicians in the year 2000 and calculate percentages
party_comp_2000 <- affiliations %>%
  inner_join(mandates, by = c("MANDATE_ID" = "ID")) %>%
  filter(MANDATE_START_YEAR <= 2000, MANDATE_END_YEAR >= 2000) %>%
  group_by(ASSEMBLY, PARTY) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(ASSEMBLY) %>%
  mutate(percentage = (count / sum(count)) * 100) %>%
  ungroup() %>%
  mutate(PARTY_GROUPED = if_else(percentage < 2, "Small Parties", PARTY))

# 2. Re-summarize by grouped parties to aggregate the small ones properly
party_comp_2000_grouped <- party_comp_2000 %>%
  group_by(ASSEMBLY, PARTY_GROUPED) %>%
  summarise(percentage = sum(percentage), count = sum(count), .groups = 'drop')

write_xlsx(party_comp_2000_grouped, "party_composition_2000_grouped.xlsx")

# 4. Create the Faceted Pie Chart (using coord_polar)
party_comp_2000_grouped %>%
  ggplot(aes(x = "", y = percentage, fill = PARTY_GROUPED)) +
  geom_col(width = 1, color = "white") +
  coord_polar("y", start = 0) +
  facet_wrap(~ASSEMBLY) +
  labs(
    title = "Party Composition of Assemblies in the Year 2000",
    subtitle = "Percentage distribution (parties < 2% grouped into 'Small Parties')",
    x = "",
    y = "",
    fill = "Party"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5),
    legend.position = "right"
  )
```

### Control of Major parties. Across Cantonal and Executive Council in political representations seems concentrated in few major parties like SVP, SP and FDP, in which the three together make the majority active mandates



## Part 4. Historical Evolution of Party affiliations 


```{r}
library(purrr)
library(tidyr)
library(dplyr)
library(ggplot2)

# 1. Expand mandates and join safely with affiliations using relationship = "many-to-many"
historical_parties <- mandates %>%
  filter(!is.na(MANDATE_START_YEAR), !is.na(MANDATE_END_YEAR), MANDATE_START_YEAR <= MANDATE_END_YEAR) %>%
  mutate(years = map2(MANDATE_START_YEAR, MANDATE_END_YEAR, seq)) %>%
  unnest(years) %>%
  filter(years >= 1980, years <= 2015) %>% # Limited to 2015
  inner_join(affiliations, by = c("ID" = "MANDATE_ID"), relationship = "many-to-many")

# 2. Calculate yearly percentages and group small parties (< 2%)
party_evolution_history <- historical_parties %>%
  group_by(ASSEMBLY, years, PARTY) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(ASSEMBLY, years) %>%
  mutate(percentage = (count / sum(count)) * 100) %>%
  ungroup() %>%
  mutate(PARTY_GROUPED = if_else(percentage < 2, "Small Parties", PARTY)) %>%
  group_by(ASSEMBLY, years, PARTY_GROUPED) %>%
  summarise(percentage = sum(percentage), .groups = 'drop')

# 3. Create clean line plot
party_evolution_history %>%
  ggplot(aes(x = years, y = percentage, color = PARTY_GROUPED)) +
  geom_line(linewidth = 1) +
  facet_wrap(~ASSEMBLY, scales = "free_y") +
  labs(
    title = "Evolution of Party Affiliations (1980 - 2015)",
    subtitle = "Tracking percentage shifts and balance of power among major parties",
    x = "Year",
    y = "Percentage of Active Politicians (%)",
    color = "Party"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )
```

### Evolving Power within the Cantonal Council. From 1980 to 2015 Cantonal Council shows a clear realignment among major political forces, FDP shows dominant in 1980s and early 990s, SVP appears between 1990s and 2000s, becoming one of the leading parties along with SP 


# Analyzing the Lifespans of Politicians


## Part 5. Politicians Lifespans and Title Analysis 


```{r}
library(dplyr)
library(ggplot2)
library(writexl)

# 1. Calculate lifespan, create title indicator, and filter realistic age ranges
persons_lifespan <- persons %>%
  filter(!is.na(YEAR_OF_DEATH), !is.na(YEAR_OF_BIRTH)) %>%
  mutate(
    LIFESPAN = YEAR_OF_DEATH - YEAR_OF_BIRTH,
    HAS_TITLE = if_else(!is.na(TITLE) & TITLE != "", "With Title", "Without Title")
  ) %>%
  filter(LIFESPAN > 18, LIFESPAN < 110) # Filtrar valores atípicos lógicos

# 2. Compute statistical summary (mean, sd, count)
lifespan_stats <- persons_lifespan %>%
  group_by(HAS_TITLE) %>%
  summarise(
    mean_lifespan = mean(LIFESPAN, na.rm = TRUE),
    sd_lifespan = sd(LIFESPAN, na.rm = TRUE),
    median_lifespan = median(LIFESPAN, na.rm = TRUE),
    count = n(),
    .groups = 'drop'
  )

print(lifespan_stats)

# 3. Run a statistical test (Welch's t-test) to compare means
t_test_result <- t.test(LIFESPAN ~ HAS_TITLE, data = persons_lifespan)
print(t_test_result)

# 4. Draw an appropriate plot (Boxplot or Violin plot with jitter) using pipes
persons_lifespan %>%
  ggplot(aes(x = HAS_TITLE, y = LIFESPAN, fill = HAS_TITLE)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.1, size = 1) +
  labs(
    title = "Comparison of Politician Lifespans by Title Status",
    subtitle = "Distribution of lifespans for politicians with and without titles",
    x = "Title Status",
    y = "Lifespan (Years)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "none"
  )

write_xlsx(lifespan_stats, "lifespan_title_analysis.xlsx")
```

### VISUAL IMPRESSION: The boxplot shows strong overlap between groups, suggesting only a small apparent difference in longevity.

### Statistical vs. Visual: 
The boxplot suggested both groups were nearly identical, but the Welch t-test revealed a highly significant 5.5‑year longevity gap—showing how dense data can visually mask real differences.”


## Part 6. Lifespan Analysis before and after 1918 by birth year


```{r}
library(dplyr)
library(ggplot2)

# 1. Prepare data, create lifespan, title indicator, and birth cohort groups via pipes
persons_cohorts <- persons %>%
  filter(!is.na(YEAR_OF_DEATH), !is.na(YEAR_OF_BIRTH)) %>%
  mutate(
    LIFESPAN = YEAR_OF_DEATH - YEAR_OF_BIRTH,
    HAS_TITLE = if_else(!is.na(TITLE) & TITLE != "", "With Title", "Without Title"),
    BIRTH_GROUP = if_else(YEAR_OF_BIRTH < 1918, "Born Before 1918", "Born 1918 or After")
  ) %>%
  filter(LIFESPAN > 18, LIFESPAN < 110)

# 2. Run separate t-test for the "Born Before 1918" group
cat("=== T-TEST: POLITICIANS BORN BEFORE 1918 ===\n")
t_test_before <- persons_cohorts %>%
  filter(BIRTH_GROUP == "Born Before 1918") %>%
  { t.test(LIFESPAN ~ HAS_TITLE, data = .) }
print(t_test_before)

# 3. Run separate t-test for the "Born 1918 or After" group
cat("\n=== T-TEST: POLITICIANS BORN 1918 OR AFTER ===\n")
t_test_after <- persons_cohorts %>%
  filter(BIRTH_GROUP == "Born 1918 or After") %>%
  { t.test(LIFESPAN ~ HAS_TITLE, data = .) }
print(t_test_after)

# 4. Create the Facet Chart using pipes
persons_cohorts %>%
  ggplot(aes(x = HAS_TITLE, y = LIFESPAN, fill = HAS_TITLE)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.1, size = 1) +
  facet_wrap(~BIRTH_GROUP) +
  labs(
    title = "Lifespan Comparison by Title Status and Birth Cohort",
    subtitle = "Comparing politicians born before 1918 vs. on or after 1918",
    x = "Title Status",
    y = "Lifespan (Years)",
    fill = "Title Status"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "none"
  )
```

### Interpretation of born before 1918: Title status appears linked to longer lifespan, likely reflecting historical class and resource differences
### Interpretation of born in 1918 or after: The longevity gap completely disappears in the younger generation.

### Statistical vs Visual interpretation: 
Before 1918, titled individuals lived significantly longer; after 1918, the difference disappeared, reflecting modern equalization in health and living conditions


## Part 7. Number of Mandates of the top 10 Politicians

```{r}
# ==============================================================================
# PART 7: TOP 10 POLITICIANS WITH THE MOST MANDATES
# ==============================================================================

library(dplyr)
library(ggplot2)

# 1. Calculate mandates per politician, take Top 10, and join with persons
top_10_politicians <- mandates %>%
  group_by(PERSON_ID) %>%
  summarise(mandate_count = n(), .groups = 'drop') %>%
  arrange(desc(mandate_count)) %>%
  slice_head(n = 10) %>%
  left_join(persons, by = c("PERSON_ID" = "ID")) %>%
  mutate(full_name = paste(FIRSTNAME, LASTNAME))

# 2. Display as a horizontal bar chart
top_10_politicians %>%
  ggplot(aes(x = reorder(full_name, mandate_count), y = mandate_count, fill = mandate_count)) +
  geom_col(show.legend = FALSE) +
  coord_flip() + 
  labs(
    title = "Top 10 Politicians by Number of Mandates",
    subtitle = "Politicians with the highest historical activity across assemblies",
    x = "Politician Name",
    y = "Number of Mandates"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.y = element_text(size = 10)
  )
```

## Part 8. Simultaneous Mandates of te top 10 politicians


```{r}
library(dplyr)
library(ggplot2)

# 1. Ensure clean mandate data with valid start and end years
mandates_clean <- mandates %>%
  filter(!is.na(MANDATE_START_YEAR), !is.na(MANDATE_END_YEAR), 
         MANDATE_START_YEAR <= MANDATE_END_YEAR)

# 2. Find overlaps using a self-join by PERSON_ID
# Two mandates (i and j) for the same person overlap if: 
# START_i <= END_j AND END_i >= START_j (excluding the exact same mandate record)
overlapping_mandates <- mandates_clean %>%
  inner_join(mandates_clean, by = "PERSON_ID", relationship = "many-to-many") %>%
  filter(ID.x != ID.y) %>% 
  filter(MANDATE_START_YEAR.x <= MANDATE_END_YEAR.y & MANDATE_END_YEAR.x >= MANDATE_START_YEAR.y)

# 3. Count how many overlapping mandate instances each politician has, and take the Top 10
top_overlapping <- overlapping_mandates %>%
  group_by(PERSON_ID) %>%
  summarise(overlap_count = n(), .groups = 'drop') %>%
  arrange(desc(overlap_count)) %>%
  slice_head(n = 10) %>%
  left_join(persons, by = c("PERSON_ID" = "ID")) %>%
  mutate(full_name = paste(FIRSTNAME, LASTNAME))

# 4. Plot the Top 10 using a horizontal bar chart
top_overlapping %>%
  ggplot(aes(x = reorder(full_name, overlap_count), y = overlap_count, fill = overlap_count)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 10 Politicians with Simultaneous Mandates",
    subtitle = "Politicians holding multiple overlapping mandates across historical periods",
    x = "Politician Name",
    y = "Number of Overlapping Mandate Instances"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.y = element_text(size = 10)
  )
```

## Part 9. Top 10 politicians with multiple party affiliations 


```{r}
library(dplyr)
library(ggplot2)

# 1. Join mandates and affiliations using PERSON_ID with explicit many-to-many relationship
politician_parties <- mandates %>%
  inner_join(affiliations, by = c("PERSON_ID" = "PERSON_ID"), relationship = "many-to-many") %>%
  filter(!is.na(PARTY), PARTY != "")

# 2. Count how many distinct parties each politician has been affiliated with
party_switchers <- politician_parties %>%
  group_by(PERSON_ID) %>%
  summarise(distinct_parties = n_distinct(PARTY), .groups = 'drop') %>%
  filter(distinct_parties > 1) %>% 
  arrange(desc(distinct_parties)) %>%
  slice_head(n = 10) %>%
  left_join(persons, by = c("PERSON_ID" = "ID")) %>%
  mutate(full_name = paste(FIRSTNAME, LASTNAME))

# 3. Plot the Top 10 Party Switchers using a horizontal bar chart
party_switchers %>%
  ggplot(aes(x = reorder(full_name, distinct_parties), y = distinct_parties, fill = distinct_parties)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Top 10 Politicians Who Changed Party Affiliations",
    subtitle = "Politicians with the highest number of distinct party affiliations over time",
    x = "Politician Name",
    y = "Number of Distinct Parties"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.y = element_text(size = 10)
  )
```


## Part 10. Mapping Politicians


```{r}
library(dplyr)
library(tidygeocoder)
library(leaflet)

# 1. Join persons with addresses using the CITY field and take a sample of 20
sample_politicians <- persons %>%
  inner_join(addresses, by = c("ID" = "PERSON_ID")) %>% 
  filter(!is.na(CITY), CITY != "") %>%
  slice_sample(n = 20) %>%
  mutate(full_address = paste(CITY, "Switzerland")) 

# 2. Convert the city names into geographic coordinates (latitude and longitude)
geocoded_data <- sample_politicians %>%
  geocode(full_address, method = 'osm', lat = latitude, long = longitude)

# 3. Interactive map with leaflet 
geocoded_data %>%
  leaflet() %>%
  addTiles() %>% 
  addMarkers(
    ~longitude, ~latitude, 
    popup = ~paste(FIRSTNAME, LASTNAME, "<br>City:", CITY)
  )
```


