---
title: "FDS Final Project: Report #1"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(scipen = 999) 
# Avoid notation
```

```{r}
library(tidyverse)
library(rvest)
library(janitor)
library(knitr)
library(httr)
library(ggplot2)
library(leaflet)
library(tidygeocoder)
library(patchwork)
```


# Part 1. Importing Data and Cleaning

## Import an Scraping

```{r}
# Web extraction 
base_url <- "https://epfl-exts.github.io/rental-scrape/"
main_page <- read_html(base_url)

raw_rentals <- tibble(
  location = main_page %>% 
    html_nodes(".address p") %>% 
    html_text(trim = TRUE),
  price = main_page %>%
    html_nodes(".price strong") %>%
    html_text(trim = TRUE),
  object_type = main_page %>%
    html_nodes(".object-type") %>%
    html_text(trim = TRUE),
  rooms = main_page %>%
    html_nodes(".rooms strong") %>%
    html_text(trim = TRUE),
  living_space = main_page %>%
    html_nodes(".living-space strong") %>%
    html_text(trim = TRUE),
  floor = main_page %>%
    html_nodes(".floor strong") %>%
    html_text(trim = TRUE),
  usable_surface = main_page %>%
    html_nodes(".usable-surface strong") %>%
    html_text(trim = TRUE),
  availability = main_page %>%
    html_nodes(".availability strong") %>%
    html_text(trim = TRUE)
  )

# Cleaning Data 
rentals_final <- raw_rentals %>%
  mutate(price = parse_number(price),
    rooms = parse_number(rooms),
    living_space = parse_number(living_space),
    usable_surface = parse_number(usable_surface),
    floor = case_when(str_detect(tolower(floor), "underground|sous-sol") ~ -1,
      str_detect(tolower(floor), "ground|rez|rdc|planta baja") ~ 0,
      TRUE ~ suppressWarnings(as.numeric(parse_number(floor)))),
    location = if_else(is.na(location) | location == "", "Sur demande", as.character(location)),
    is_hidden = if_else(str_detect(tolower(location), "on request|sur demande"), "On Request", "Available"))

rentals_final %>%
  summary() %>%
  kable(caption = "Summary of Cleaned Data")

write_csv(rentals_final, "rental_data_final_epfl_612.csv")
```

**Data Cleaning and Preparation**
_The cleaned data-set ends up with 612 rows and 9 variables , and saved as rental_data_final_epfl_612.csv for further analysis._

**Explanation parse_number():** _I used it to clean numeric strings that contain suffixes or symbols. To prevent extreme distortions caused by outliers in price_m2 often produced by incorrectly recorded very small surface values, we filter out infinite values, NAs, and keep only logically coherent ranges of living area._
**parse_number()** _It's use to find the first numbers inside of a chain of text, ignoring characters or other symbols, spaces or commas_



# Exploratory Data Analysis

## Part 2. Price vs. Living Space

```{r}
rentals_final %>%
  mutate(postcode = str_extract(location, "\\d{4}")) %>%
  filter(!is.na(postcode)) %>%
  count(postcode) %>%
  ggplot(aes(x = n, y = reorder(postcode, n))) +
  geom_col(fill = "steelblue") +
  labs(title = "Properties by Postcode",
       x = "Number of Properties",
       y = "Postcode") +
  theme_minimal()

# Correction
rentals_final_postal <- rentals_final %>%
  mutate(postcode = str_extract(location, "\\d{4}")) %>%
  filter(floor %in% 1:6)
```

**INTERPRETATION:**  _As the living space increase, the price as well rise up, normal, bigger the space of living more expensive became._


## Part 3. Number of properties by postcode

```{r}
rentals_final %>%
  mutate(postcode = str_extract(location, "\\d{4}")) %>%
  count(postcode) %>%
  ggplot(aes(x = postcode, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Properties by Postcode",
    x = "Postcode",
    y = "Number of Properties") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**Comments:** _The plot shows that the agency is more active in other areas where there is more listings, other areas have less listings indicating lower activity_


## Part 4. Price vs. Living Space by Postcode and Floor 

```{r}

#Floor 1 to 6
rentals_final_postal <- rentals_final %>%
  mutate(postcode = str_extract(location, "\\d{4}")) %>%
  filter(floor %in% 1:6) 

combo_summary <- rentals_final_postal %>%
  group_by(postcode, floor) %>%
  summarise(mean_price = mean(price, na.rm = TRUE), .groups = "drop")

most_expensive_combo <- combo_summary %>%
  slice_max(mean_price, n = 1)
least_expensive_combo <- combo_summary %>%
  slice_min(mean_price, n = 1)

rentals_final_postal %>%
  ggplot(aes(x = living_space, y = price, color = postcode)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ floor) +
  labs(title = "Price vs Living Space by Postcode and Floor (Floors 1-6)",
       x = "Living Space (m²)",
       y = "Price (CHF)",
       color = "Postcode") +
  theme_minimal()
```

The most expensive combination is postcode `r most_expensive_combo$postcode` on floor
`r most_expensive_combo$floor` with an average price of `r round(most_expensive_combo$mean_price, 2)` CHF

The least expensive combination is postcode `r least_expensive_combo$postcode` on floor
`r least_expensive_combo$floor` with an average of `r round(least_expensive_combo$mean_price, 2)` CHF


# Analysis

## Part 5.

```{r}

# Prices comparison
p_price <- rentals_final %>%
  ggplot(aes(x = is_hidden, y = price, fill = is_hidden)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.2, fill = "white") +
  labs(title = "Price Comparison", x = "Status", y = "Price (CHF)") +
  theme_minimal() +
  theme(legend.position = "none")

# Living space comparison 
p_space <- rentals_final %>%
  ggplot(aes(x = is_hidden, y = living_space, fill = is_hidden)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.2, fill = "white") +
  labs(title = "Living Space Comparison", x = "Status", y = "Living Space (m²)") +
  theme_minimal() +
  theme(legend.position = "none")

# Floor comparison 
p_floor <- rentals_final %>%
  filter(!is.na(floor)) %>%
  ggplot(aes(x = is_hidden, y = as.factor(floor), fill = is_hidden)) +
  geom_boxplot(width = 0.4) +
  labs(title = "Floor Comparison", x = "Status", y = "Floor") +
  theme_minimal() +
  theme(legend.position = "none")


p_price / p_space / p_floor
```

_COMMENTS_

**Price Comparison:** _The violin and box plots indicate that "On request" properties tend to have higher overall price levels and a broader upper tail distribution, suggesting more expensive and high end listings compared to available properties._
**Living Space Comparison:** _"On Request properties also tend to have larger living areas, suggesting that listings with hidden prices are typically bigger homes or apartments._
**Floor Comparison:** _The floor distribution shows that both "Available" and "On Request" properties appear across similar floor levels, with no clear bias toward any specific floors in either group._


## Part 6. Statistical Summary and Hypothesis Testing for Price for Square Meter


```{r}
rentals_stats <- rentals_final %>%
  mutate(price_m2 = price / living_space) %>%
  filter(!is.na(price_m2) & !is.infinite(price_m2) & living_space > 5)

# Summary table
rentals_stats %>%
  group_by(is_hidden) %>%
  summarise(count = n(),
    median = median(price_m2, na.rm = TRUE),
    mean = mean(price_m2, na.rm = TRUE),
    sd = sd(price_m2, na.rm = TRUE),
    min = min(price_m2, na.rm = TRUE),
    max = max(price_m2, na.rm = TRUE)) %>%
  kable(caption = "Statistical Summary of Price per m² by Property Status")
```

**COMMENTS:** _" On Request" properties show higher average and median price per square meter, along with a wider spread in values. The larger standard deviation, minimum, and maximum suggest greater variability and the presence of more high end listings in this group._



## Part 6.1 T-Test

```{r}
t_test_m2 <- t.test(price_m2 ~ is_hidden, data = rentals_stats)
print(t_test_m2)
```

**CONCLUSION:** _The t‑test shows no statistically significant differece in average price per square meter between "Available" and "On Request" properties. The small numerical gap seen in the descriptive summary is not meaningful and likely reflects normal sample variation rather than a real difference in unit pricing._



## Part 7. Analysis using Total Price (CHF)


```{r}
rentals_ready <- rentals_final 

rentals_ready %>%
  group_by(is_hidden) %>%
  summarise(count = n(),
    median = median(price, na.rm = TRUE),
    mean = mean(price, na.rm = TRUE),
    sd = sd(price, na.rm = TRUE),
    min = min(price, na.rm = TRUE),
    max = max(price, na.rm = TRUE)) %>%
  kable(caption = "Statistical Summary of Total Price by Property Status")
```

**INTERPRETATION:** _"On Request" properties shows a higher average and median total price than "Available"  ones (about 4,030 CHF vs. 3,449 CHF, and 3,500 CHF vs. 2,767.5 CHF). This pattern suggests a right skewed distribution where hidden or un priced listings tend to fall into higher overall price ranges._



## Part 7.1 T-Test


```{r}
t_test_total <- t.test(price ~ is_hidden, data = rentals_ready)
print(t_test_total)
```

**OBSERVATION AND CONCLUSIONS:** _The t-test shows that "On Request" properties have a significantly higher average total price than "Available" ones. This points to a clear price gap between publicly listed properties and those with hidden or unlisted prices._



# Part 8. Geo-coding and Mapping with {tidygeocoder} and {leaflet}

```{r}
# 30 Properties
map_sample <- rentals_ready %>%
  filter(!is.na(location) & location != "Sur demande") %>%
  sample_n(min(30, nrow(.))) %>%
  geocode(address = location, method = "arcgis", lat = latitude, lon = longitude)

# Map
map_sample %>%
  filter(!is.na(latitude) & !is.na(longitude)) %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(lng = ~longitude, 
    lat = ~latitude,
    popup = ~paste("<b>Tipo:</b>", object_type, "<br>",
      "<b>Precio:</b> CHF", price, "<br>",
      "<b>Ubicación:</b>", location))
```


