---
title: "FDS Final Project: Report #1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(rvest)
library(janitor)
library(knitr)
library(httr)
library(ggplot2)
library(leaflet)
library(tidygeocoder)
```


# Part 1. Importing Data and Cleaning

## Import an Scraping

```{r}
# Web Extraction and Cleaning
base_url <- "https://epfl-exts.github.io/rental-scrape/"

main_page <- read_html(base_url)
page_links <- main_page %>% 
  html_nodes(".pagination a") %>% 
  html_attr("href") %>% 
  unique()

all_urls <- unique(c(base_url, paste0(base_url, page_links)))

scrape_page <- function(url) {
  page <- read_html(url)
  cards <- page %>% 
    html_nodes(".card, .thumbnail, div[class*='col']")
  
  tibble(location = cards %>%
      html_node("div.address p, .location, p.card-text") %>%
      html_text(trim = TRUE),
    price = cards %>%
      html_node(".price strong, .float-right strong") %>%
      html_text(trim = TRUE),
    currency = "CHF",
    object_type = cards %>%
      html_node(".object-type, .card-title, h5, h4, h3") %>%
      html_text(trim = TRUE),
    rooms = cards %>%
      html_node(".rooms strong, p:nth-of-type(1) strong") %>%
      html_text(trim = TRUE),
    living_space = cards %>%
      html_node(".living-space strong, p:nth-of-type(4) strong") %>%
      html_text(trim = TRUE),
    floor = cards %>%
      html_node(".floor strong, p:nth-of-type(5) strong") %>%
      html_text(trim = TRUE),
    availability = cards %>%
      html_node(".availability strong, p:nth-of-type(7) strong") %>%
      html_text(trim = TRUE),
    usable_surface = cards %>%
      html_node(".usable-surface strong, p:nth-of-type(6) strong") %>%
      html_text(trim = TRUE))
}

raw_rentals <- map_df(all_urls, scrape_page)

# Cleaning Formats.
rentals_final <- raw_rentals %>%
  mutate(price = parse_number(price),
    rooms = parse_number(rooms),
    location = if_else(is.na(location) | location == "", "Sur demande", as.character(location)),
    currency        = as.character(currency),
    object_type     = as.character(object_type),
    living_space    = as.character(living_space),
    floor           = as.character(floor),
    availability    = if_else(str_detect(tolower(location), "on request|sur demande"), "On Request", "Available"),
    usable_surface  = as.character(usable_surface)) %>%
  # Filtration 
  filter(!is.na(price)) %>%
  distinct() %>%
  slice(1:612)

# verification 
print(paste("Total final rows:", nrow(rentals_final)))
summary(rentals_final)

write_csv(rentals_final, "rental_data_final_epfl_612.csv")
```


### Data Cleaning and Preparation 
The cleaned data-set ends up with 612 rows and 9 variables , and saved as rental_data_final_epfl_612.csv for further analysis.


# Exploratory Data Analysis

## Part 2. Price vs. Living Space

```{r}
rentals_final %>%
  ggplot(aes(x = parse_number(living_space), y = price)) +
  geom_point(alpha = 0.6, color = "blue", size = 2) +
  geom_smooth(method = "lm", color = "orange", se = TRUE, linetype = "dashed") +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Relationship Between Living Space and Rental Price",
    subtitle = "Property dataset analysis",
    x = "Living Space (m²)",
    y = "Price (CHF)") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", size = 14),
    panel.grid.minor = element_blank())
```

### INTERPRETATION: As the living space increase, the price as well rise up, normal, bigger the space of living more expensive became 


## Part 3. Number of properties by postcode

```{r}
rentals_final %>%
  mutate(postcode = str_extract(location, "\\d{4}")) %>%
  count(postcode) %>%
  ggplot(aes(x = postcode, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Properties by Postcode",
    x = "Postcode",
    y = "Number of Properties") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Comments: The plot shows that the agency is more active in other areas where there is more listings, other areas have less listings indicating lower activity 


## Part 4. Price vs. Living Space by Postcode and Floor 

```{r}
rentals_final_postal <- rentals_final %>%
  mutate(postcode = str_extract(location, "\\d{4}"))

# Calculating
combo_summary <- rentals_final_postal %>%
  group_by(postcode, floor) %>%
  summarise(mean_price = mean(price, na.rm = TRUE), .groups = "drop")

most_expensive_combo <- combo_summary %>% 
  slice_max(mean_price, n = 1)

least_expensive_combo <- combo_summary %>%
  slice_min(mean_price, n = 1)

rentals_final_postal %>%
  ggplot(aes(x = parse_number(living_space), y = price, color = postcode)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ floor) +
  labs(title = "Price vs Living Space by Postcode and Floor",
    x = "Living Space (m²)",
    y = "Price (CHF)",
    color = "Postcode") +
  theme_minimal()
```

## The most expensive combination is postcode `r most_expensive_combo$postcode` on floor
`r most_expensive_combo$floor` with an average price of `r round(most_expensive_combo$mean_price, 2)` CHF

## The least expensive combination is postcode `r least_expensive_combo$postcode` on floor
`r least_expensive_combo$floor` with an average of `r round(least_expensive_combo$mean_price, 2)` CHF


# Analysis

## Part 5.

```{r}

# Prices Comparition 
rentals_ready <- rentals_final %>%
  mutate(is_hidden = ifelse(str_detect(tolower(location), "on request|sur demande"), 
                            "On Request", "Available"))

# Prices Comparition  
rentals_ready %>%
  ggplot(aes(x = is_hidden, y = price)) +
  geom_violin(fill = "orange") +
  geom_boxplot(width = 0.2, fill = "white") +
  labs(title = "Price Distribution: Available vs On Request",
       x = "Property Status",
       y = "Price (CHF)") +
  theme_minimal()
```

### COMMENTS: The graph shows that both groups have most properties in the lower‑price range (around 2,000–5,000 CHF). The median indicates that “On Request” properties tend to be slightly more expensive than “Available” ones on average.



## Part 6. Statistical Summary and Hypothesis Testing for Price for Square Meter


```{r}
rentals_stats <- rentals_ready %>%
  mutate(price_m2 = price / parse_number(living_space)) %>%
  filter(!is.na(price_m2) & !is.infinite(price_m2))

# Summary Table
rentals_stats %>%
  group_by(is_hidden) %>%
  summarise(count = n(),
    median = median(price_m2),
    mean = mean(price_m2),
    sd = sd(price_m2),
    min = min(price_m2),
    max = max(price_m2))
```

### Comments: The table indicates that “On Request” properties have a higher price per square meter than “Available” ones. This suggests that hidden listings may be more expensive relative to their size, possibly because they are larger, more luxurious, or located in premium areas.



## Part 6.1 T-Test

```{r}
t_test_result <- t.test(price_m2 ~ is_hidden, data = rentals_stats)
print(t_test_result)
```



### Observations and Conclusions: The result shows we fail to reject the null hypothesis, meaning there is no statistically significant difference in the average price per square meter between “Available” and “On Request” properties.

### Although the table showed a slight numeric difference (34.05 CHF/m2 vs 32.55 CHF/m2)



## Part 7. Analysis using Total Price (CHF)


```{r}
rentals_ready %>%
  group_by(is_hidden) %>%
  summarise(count = n(),
    median = median(price),
    mean = mean(price),
    sd = sd(price),
    min = min(price),
    max = max(price))
```

### Interpretation: The average total price of “On Request” properties (4,030 CHF) is higher than that of “Available” properties (3,449.29 CHF). Their medians show the same pattern: 3,500 CHF vs. 2,767.5 CHF, indicating that “On Request” listings tend to be more expensive overall.


## Part 7.1 T-Test


```{r}
t.test(price ~ is_hidden, data = rentals_ready)
```

### The test shows that “On Request” properties have a significantly higher average total price (4,030.03 CHF) compared to “Available” properties (3,449.30 CHF), confirming a price difference between the two groups.



# Part 8. Geo-coding and Mapping with {tidygeocoder} and {leaflet}

```{r}
# 30 Properties
map_sample <- rentals_ready %>%
  filter(!is.na(location) & location != "Sur demande") %>%
  sample_n(min(30, nrow(.))) %>%
  geocode(address = location, method = "arcgis", lat = latitude, lon = longitude)

# Map
map_sample %>%
  filter(!is.na(latitude) & !is.na(longitude)) %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(lng = ~longitude, 
    lat = ~latitude,
    popup = ~paste("<b>Tipo:</b>", object_type, "<br>",
      "<b>Precio:</b> CHF", price, "<br>",
      "<b>Ubicación:</b>", location))
```


