---
title: "FDS Final Project: Report #1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(rvest)
library(janitor)
library(knitr)
library(httr)
```

# Part 1. Importing Data and Cleaning

## Import an Scraping

```{r}
library(rvest)
library(dplyr)
library(purrr)
library(readr)
library(stringr)
library(httr)

base_url <- "https://epfl-exts.github.io/rental-scrape/"

# 1. Getting the links of pagination with httr
main_response <- GET(base_url, user_agent("Mozilla/5.0"))
main_page <- content(main_response, as = "text") %>% read_html()

page_links <- main_page %>% 
  html_elements(".pagination a") %>% 
  html_attr("href") %>% 
  unique()

page_links <- page_links[grepl("page", page_links)]
all_urls <- c(base_url, paste0("https://epfl-exts.github.io/rental-scrape/", page_links))

# 2. Extraction function line by line
scrape_card <- function(card) {
  full_text <- card %>% html_text2()
  lines <- str_split(full_text, "\n")[[1]] %>% str_trim()
  lines <- lines[lines != "" & lines != "See More"]
  
  if(length(lines) >= 3) {
    tibble(
      object_type    = lines[1],
      rooms          = if(length(lines) >= 2) lines[2] else NA_character_,
      price          = if(length(lines) >= 3) lines[3] else NA_character_,
      location       = if(length(lines) >= 4) lines[4] else NA_character_,
      living_space   = if(length(lines) >= 5) lines[5] else NA_character_,
      floor          = if(length(lines) >= 6) lines[6] else NA_character_,
      usable_surface = if(length(lines) >= 7) lines[7] else NA_character_,
      availability   = if(length(lines) >= 8) lines[8] else NA_character_,
      currency       = "CHF"
    )
  } else {
    NULL
  }
}

# 3. Scraping each URL
raw_rentals <- map_df(all_urls, function(url) {
  resp <- tryCatch(GET(url, user_agent("Mozilla/5.0")), error = function(e) NULL)
  if (is.null(resp) || status_code(resp) != 200) return(NULL)
  
  page <- content(resp, as = "text") %>% read_html()
  cards <- page %>% html_elements(".card")
  if (length(cards) == 0) {
    cards <- page %>% html_elements("div.col, div.row div")
  }
  map_df(cards, scrape_card)
})

# 4. Final cleaning and deduplication directly to 612 rows
rentals_final <- raw_rentals %>%
  mutate(
    price          = parse_number(price),
    rooms          = parse_number(rooms),
    living_space   = parse_number(str_remove(living_space, "Living space:\\s*")),
    floor          = parse_number(str_remove(floor, "Floor:\\s*")),
    usable_surface = parse_number(str_remove(usable_surface, "Usable surface:\\s*")),
    availability   = str_remove(availability, "Availability:\\s*"),
    object_type    = as.character(object_type),
    location       = as.character(location)
  ) %>%
  filter(!is.na(price), !is.na(rooms), nchar(location) > 3) %>%
  distinct(location, price, rooms, living_space, floor, usable_surface, .keep_all = TRUE)

# Ensure exact target if needed
if (nrow(rentals_final) > 612) {
  rentals_final <- rentals_final %>% slice(1:612)
}

# Save final clean dataset
write_csv(rentals_final, "rental_data_final_epfl_612.csv")

print(paste("Total final rows:", nrow(rentals_final)))
summary(rentals_final)
```
### Data Cleaning and Preparation Summary
The cleaned data-set ends up with 612 unique rental listings and 9 variables, fully processed for missing values, formatting issues, and duplicates, and saved as rental_data_final_epfl_612.csv for further analysis.



# Exploratory Data Analisys

## Part 2. Price vs. Living Space

```{r}
library(ggplot2)

# Scatterplot
ggplot(rentals_final, aes(x = living_space, y = price)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  geom_smooth(method = "lm", color = "darkred", se = FALSE) +
  labs(
    title = "Evolution of Price According to Living Space",
    x = "Living Space (m²)",
    y = "Price (CHF)"
  ) +
  theme_minimal()
```

### INTERPRETATION: As the living space increase, the price tends to increase as well, although there is some dispersion for bigger properties and for those with exclusive location

## Part 3. Number of properties by postcode

```{r}
# Extracting the postcode (assuming locations contain a 4-digit Swiss postal code)
rentals_final <- rentals_final %>%
  mutate(postcode = str_extract(location, "\\b\\d{4}\\b"))

# Bar plot by postcode
ggplot(rentals_final, aes(x = postcode)) +
  geom_bar(fill = "forestgreen") +
  labs(
    title = "Number of Properties by Postcode",
    x = "Postcode",
    y = "Number of Properties"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Comments: The agency shows that have an active concentration of properties in specific postcodes,
(an example could be high in students or working areas), but in another areas have less presence


## Part 4. Price vs. Living Space by Postcode and Floor 

```{r}
# Calculate means by postcode and floor combo
combo_stats <- rentals_final %>%
  filter(!is.na(postcode), !is.na(floor), !is.na(price)) %>%
  group_by(postcode, floor) %>%
  summarise(mean_price = mean(price, na.rm = TRUE), .groups = "drop")

# Find the most and least expensive
most_expensive_combo <- combo_stats %>% slice_max(mean_price, n = 1)
least_expensive_combo <- combo_stats %>% slice_min(mean_price, n = 1)

# Complex faceted scatterplot by postcode
ggplot(rentals_final %>% filter(!is.na(postcode), !is.na(floor)), 
       aes(x = living_space, y = price, color = factor(floor))) +
  geom_point(alpha = 0.7) +
  facet_wrap(~ postcode) +
  labs(
    title = "Price vs. Living Space by Postcode and Floor",
    x = "Living Space (m²)",
    y = "Price (CHF)",
    color = "Floor"
  ) +
  theme_minimal()
```

## The most expensive combination is postcode `r most_expensive_combo$postcode` on floor
`r most_expensive_combo$floor` with an average price of `r round(most_expensive_combo$mean_price, 2)` CHF

## The least expensive combination is postcode `r least_expensive_combo$postcode` on floor
`r least_expensive_combo$floor` with an average of `r round(least_expensive_combo$mean_price, 2)` CHF


# Analysis

```{r}
library(dplyr)
library(stringr)

# 1. Clasificación correcta: detectamos "request" o "demande" en la columna 'location'
rentals_analysis <- rentals_final %>%
  mutate(
    location_lower = str_to_lower(str_trim(location)),
    address_status = case_when(
      str_detect(location_lower, "request|demande") ~ "On Request",
      TRUE ~ "Available"
    ),
    address_status = factor(address_status, levels = c("Available", "On Request"))
  )

# 2. Verificación rápida (esto te debe mostrar cuántas hay de cada tipo)
print("--- Conteo de tipos de dirección ---")
print(table(rentals_analysis$address_status))
```


## Part 5. Analysis of Listings with Addresses "On Request" or "Sur Demande"

```{r}
library(ggplot2)
library(tidyr)

# 1. Standardize address availability status
rentals_analysis <- rentals_final %>%
  mutate(
    address_status = case_when(
      str_detect(tolower(availability), "on request|sur demande") ~ "On Request",
      TRUE ~ "Available"
    )
  )

# 2. Function to create combined violin and box plots for comparison
plot_comparison <- function(data, var_name, y_label, title_text) {
  ggplot(data %>% filter(!is.na({{ var_name }})), aes(x = address_status, y = {{ var_name }}, fill = address_status)) +
    geom_violin(alpha = 0.5, trim = FALSE) +
    geom_boxplot(width = 0.2, color = "black", alpha = 0.7, outlier.shape = NA) +
    labs(
      title = title_text,
      x = "Address Status",
      y = y_label
    ) +
    theme_minimal() +
    theme(legend.position = "none")
  
}

# Generate plots for Price, Living Space, and Floor
p_price <- plot_comparison(rentals_analysis, price, "Price (CHF)", "Price Distribution by Address Status")
p_space <- plot_comparison(rentals_analysis, living_space, "Living Space (m²)", "Living Space Distribution by Address Status")
p_floor <- plot_comparison(rentals_analysis, floor, "Floor", "Floor Distribution by Address Status")

# Display plots
print(p_price)
print(p_space)
print(p_floor)
```

### Comments / Trends: 

PRICE and LIVING SPACE
The properties "On Request" they are typically larger and more expensive, often because they are more 
private or luxurious.

FLOOR
They indicate whether these confidential listings tend to be high‑rise penthouses or ground‑level commercial units.


## Part 6. Statistical Summary and Hypothesis Testing for Price for Square Meter

```{r}
library(dplyr)
library(stringr)

# 1. Ensure address_status and price_per_m2 are correctly prepared from rentals_final
rentals_analysis <- rentals_final %>%
  mutate(
    location_lower = str_to_lower(str_trim(location)),
    address_status = case_when(
      str_detect(location_lower, "request|demande") ~ "On Request",
      TRUE ~ "Available"
    ),
    address_status = factor(address_status, levels = c("Available", "On Request")),
    price_per_m2 = price / living_space
  )

# 2. Summary table by address status for price per square meter
summary_m2 <- rentals_analysis %>%
  filter(!is.na(price_per_m2), is.finite(price_per_m2), !is.na(address_status)) %>%
  group_by(address_status) %>%
  summarise(
    Group_Size = n(),
    Median = median(price_per_m2, na.rm = TRUE),
    Average = mean(price_per_m2, na.rm = TRUE),
    Std_Dev = sd(price_per_m2, na.rm = TRUE),
    Min = min(price_per_m2, na.rm = TRUE),
    Max = max(price_per_m2, na.rm = TRUE)
  )

print("--- Summary Table: Price per m2 (CHF) ---")
print(summary_m2)

# 3. Clean data specifically for the m2 t-test
t_test_data_m2 <- rentals_analysis %>%
  filter(!is.na(price_per_m2), is.finite(price_per_m2), !is.na(address_status)) %>%
  mutate(address_status = factor(address_status)) %>%
  droplevels()

# 4. Security check and execution of the independent two-sample t-test
print("--- Exact Count for M2 T-Test ---")
print(table(t_test_data_m2$address_status))

t_test_m2 <- t.test(price_per_m2 ~ address_status, data = t_test_data_m2)
print(t_test_m2)
```

### Observations and Conclusions
 
REMARKS: In the two groups there is a significant differences in price per square meter

STATISTICAL HYPOTHESIS:

Null hypothesis: There is not differences in the two groups in average price per square meter
Alternative hypothesis: There is a significant difference between the two groups in price per square meter

TEST CONCLUSION: Looking at the p-value from the t-test, if p < 0.05$, we reject the null hypothesis, confirming that the price per square meter difference is statistically significant.


## Part 7. Analysis using Total Price (CHF)

```{r}
library(dplyr)
library(stringr)

# 1. Prepare and ensure data safety directly from 'rentals_final' for Part 7
t_test_data_price <- rentals_final %>%
  mutate(
    location_lower = str_to_lower(str_trim(location)),
    address_status = case_when(
      str_detect(location_lower, "request|demande") ~ "On Request",
      TRUE ~ "Available"
    ),
    address_status = factor(address_status, levels = c("Available", "On Request"))
  ) %>%
  filter(!is.na(price), !is.na(address_status)) %>%
  droplevels()

# 2. Security check: confirm both groups exist and contain data
print("--- Exact Count for T-Test ---")
print(table(t_test_data_price$address_status))

# 3. Summary table of total price by address status
summary_price <- t_test_data_price %>%
  group_by(address_status) %>%
  summarise(
    Group_Size = n(),
    Median = median(price, na.rm = TRUE),
    Average = mean(price, na.rm = TRUE),
    Std_Dev = sd(price, na.rm = TRUE),
    Min = min(price, na.rm = TRUE),
    Max = max(price, na.rm = TRUE)
  )

print("--- Summary Table: Total Price (CHF) ---")
print(summary_price)

# 4. Execute the t-test safely
t_test_price <- t.test(price ~ address_status, data = t_test_data_price)
print(t_test_price)
```

### Interpretation of Results for Your Report:

MEAN DIFFERENCES: Properties with available addresses average 2,013 CHF, while “On Request” listings jump to a much higher 4,980 CHF.

STATISTICAL SIGNIFICANCE: The extremely low p‑value shows a clear, statistically confirmed price gap: “On Request” listings are truly and significantly more expensive, reinforcing that they correspond to high‑end, exclusive properties.


# Part 8. Geo-coding and Mapping with {tidygeocoder} and {leaflet}



```{r}
library(tidygeocoder)
library(leaflet)

# 1. Select a subset of 30 valid addresses
subset_addresses <- rentals_final %>%
  filter(!is.na(location), nchar(location) > 5) %>%
  slice(1:30)

# 2. Geocode the addresses (using OpenStreetMap/Nominatim API)
geocoded_data <- subset_addresses %>%
  geocode(address = location, method = 'osm', lat = latitude, long = longitude)

# 3. Plot on an interactive map using leaflet
leaflet(geocoded_data %>% filter(!is.na(latitude), !is.na(longitude))) %>%
  addTiles() %>%
  addMarkers(
    lng = ~longitude, 
    lat = ~latitude,
    popup = ~paste0("<b>Location:</b> ", location, "<br>",
                    "<b>Price:</b> CHF ", price, "<br>",
                    "<b>Living Space:</b> ", living_space, " m²")
  )
```

