---
title: "Meta Ads Performance and Campaign Audit"
subtitle: "Analysis of Ad Spend, Creative Fatigue, and Sales Return (ROAS)"
author: "Franklin Andrade"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 5.2,
  fig.align = "center"
)

library(tidyverse)
library(lubridate)
library(scales)
library(knitr)
library(kableExtra)

set.seed(42)
```


# Project Overview


## This report reviews the results of four advertising campaigns on Facebook and Instagram (Meta Ads) over a period of 45 days (August 1 to September 15, 2026).


The main goals are:
1. Identify when ads start losing effectiveness because audiences see them too many times (**Ad Fatigue**).
2. Measure the financial return of each ad format (**Return on Ad Spend - ROAS**).
3. Recommend how to adjust the advertising budget to get better overall results.

---

# Campaign Structure and Data


The data simulates a standard export from **Meta Ads Manager**. It includes both new audience discovery (**Prospecting**) and existing visitor campaigns (**Re-targeting**).


```{r data-generation}
# Flight dates
dates_list <- seq(as.Date("2026-08-01"), as.Date("2026-09-15"), by = "day")
total_days <- length(dates_list)

# Initial campaign setup
campaign_settings <- tribble(
  ~campaign,                    ~funnel_level,  ~objective,      ~ad_format,   ~target_group,    ~target_daily_budget,
  "New_Customer_Traffic",        "Prospecting",  "Traffic",       "Carousel",   "Broad_Public",   180,
  "Brand_Video_Views",           "Prospecting",  "Video Views",   "Reels",      "Interest_Group", 120,
  "Product_Catalog_Retargeting", "Retargeting",  "Conversions",   "Dynamic",    "Website_Visitors", 250,
  "Promo_Discount_Offers",       "Retargeting",  "Conversions",   "Single_Image","Cart_Abandoners", 150
)

# Daily campaign data generator
raw_meta_data <- campaign_settings %>%
  slice(rep(1:n(), each = total_days)) %>%
  mutate(date = rep(dates_list, times = nrow(campaign_settings))) %>%
  rowwise() %>%
  mutate(
    # Daily investment (in CHF)
    spend = max(25, round(rnorm(1, mean = target_daily_budget, sd = 15), 2)),
    
    # Cost per 1,000 views (CPM)
    average_cpm = case_when(
      target_group == "Broad_Public"     ~ rnorm(1, 14.5, 1.2),
      target_group == "Interest_Group"   ~ rnorm(1, 11.2, 0.9),
      target_group == "Website_Visitors" ~ rnorm(1, 28.5, 2.0),
      target_group == "Cart_Abandoners"  ~ rnorm(1, 36.0, 2.5)
    ),
    
    # Volume: impressions and unique people reached
    impressions = round((spend / average_cpm) * 1000),
    reach = round(impressions / runif(1, 1.10, 1.80)),
    
    # Click-through rate base
    estimated_ctr = case_when(
      ad_format == "Carousel"     ~ runif(1, 0.016, 0.024),
      ad_format == "Reels"        ~ runif(1, 0.009, 0.015),
      ad_format == "Dynamic"      ~ runif(1, 0.023, 0.038),
      ad_format == "Single_Image" ~ runif(1, 0.017, 0.028)
    ),
    clicks = round(impressions * estimated_ctr),
    
    # Conversion rate
    conversion_rate = case_when(
      objective == "Conversions" ~ runif(1, 0.032, 0.065),
      objective == "Traffic"     ~ runif(1, 0.006, 0.014),
      TRUE                       ~ runif(1, 0.001, 0.003)
    ),
    purchases = round(clicks * conversion_rate),
    
    # Average order size and revenue
    average_order = runif(1, 80, 110),
    revenue = round(purchases * average_order, 2)
  ) %>%
  ungroup()

# KPI calculations
clean_campaign_data <- raw_meta_data %>%
  mutate(
    # How many times each person saw the ad
    frequency = round(impressions / reach, 2),
    
    # Percentage of views that turned into clicks
    ctr = (clicks / impressions) * 100,
    
    # Cost per click
    cpc = ifelse(clicks > 0, spend / clicks, NA),
    
    # Return on Ad Spend: Money made divided by money spent
    roas = ifelse(spend > 0, revenue / spend, 0),
    
    # Cost per purchase
    cpa = ifelse(purchases > 0, spend / purchases, NA)
  )
```


# Key Metrics Guide

## This table explains the main metrics used in this report:


```{r metrics-table}
tribble(
  ~Metric_Name, ~What_It_Means, ~Why_It_Matters, ~Good_Benchmark,
  "Frequency", "Average times a person sees the ad", "Warns when people see the ad too often", "Keep under 3.5 per week",
  "CTR (%)", "Percentage of views that clicked", "Shows how interesting the ad is", "Above 1.5% is good",
  "CPC (CHF)", "Average cost for each click", "Measures the cost to bring people to the website", "Under CHF 1.20 for new audiences",
  "ROAS", "Revenue divided by ad spend", "Measures profit and business return", "Above 3.0x for retargeting"
) %>%
  kable(caption = "Table 1: Main Digital Marketing Metrics") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE, width = "3cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3, width = "5cm") %>%
  column_spec(4, width = "4cm")
```


# Total Performance Summary

```{r summary-table}
performance_summary <- clean_campaign_data %>%
  group_by(Campaign = campaign, Audience = funnel_level, Format = ad_format) %>%
  summarise(
    Total_Spend   = sum(spend),
    Impressions   = sum(impressions),
    Reach         = sum(reach),
    Avg_Frequency = round(mean(frequency), 2),
    Total_Clicks  = sum(clicks),
    Average_CTR   = paste0(round(mean(ctr), 2), "%"),
    Average_CPC   = round(mean(cpc, na.rm = TRUE), 2),
    Total_Sales   = sum(purchases),
    Total_Revenue = sum(revenue),
    Total_ROAS    = round(sum(revenue) / sum(spend), 2),
    .groups = "drop"
  )

performance_summary %>%
  mutate(
    Total_Spend   = dollar(Total_Spend, prefix = "CHF "),
    Total_Revenue = dollar(Total_Revenue, prefix = "CHF "),
    Average_CPC   = dollar(Average_CPC, prefix = "CHF "),
    Impressions   = comma(Impressions),
    Reach         = comma(Reach),
    Total_Clicks  = comma(Total_Clicks),
    Total_Sales   = comma(Total_Sales)
  ) %>%
  kable(caption = "Table 2: Performance Summary by Campaign (45-Day Total)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 12)
```


# Performance Charts


## Return on Ad Spend (ROAS) vs. Budget Spent


**This chart compares the money spent on each campaign against the return on sales it generated.**

```{r}
ggplot(performance_summary, aes(x = Total_Spend, y = Total_ROAS, color = Campaign)) +
  geom_point(aes(size = Total_Sales), alpha = 0.85) +
  geom_hline(yintercept = 1.0, linetype = "dashed", color = "gray50", linewidth = 0.8) +
  geom_hline(yintercept = 3.0, linetype = "dotted", color = "forestgreen", linewidth = 0.8) +
  annotate("text", x = min(performance_summary$Total_Spend) + 300, y = 1.15, 
           label = "Break-even line (ROAS = 1.0)", color = "gray40", size = 3.8) +
  annotate("text", x = min(performance_summary$Total_Spend) + 300, y = 3.15, 
           label = "Target goal (ROAS = 3.0)", color = "forestgreen", size = 3.8) +
  scale_y_continuous(labels = label_number(suffix = "x")) +
  scale_x_continuous(labels = label_dollar(prefix = "CHF ")) +
  scale_size_continuous(range = c(5, 12)) +
  labs(
    title = "Campaign Efficiency: Money Spent vs. Return (ROAS)",
    subtitle = "Larger circles represent more total sales generated",
    x = "Total Budget Spent (CHF)",
    y = "Return on Ad Spend (Revenue / Spend)",
    size = "Sales Count"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right")
```

# Ad Fatigue: What Happens When People See Ads Too Often


## When an ad is shown to the same person repeatedly, people stop paying attention and the Click-Through Rate (**CTR**) drops.


```{r fatigue-chart}
ggplot(clean_campaign_data, aes(x = frequency, y = ctr, color = ad_format)) +
  geom_point(alpha = 0.35, size = 1.8) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 1.2) +
  scale_y_continuous(labels = label_percent(scale = 1)) +
  labs(
    title = "Ad Fatigue: How Frequency Affects Click Rates",
    subtitle = "Click-Through Rate drops as users see the same ad more times",
    x = "Average Views per Person (Frequency)",
    y = "Click-Through Rate (CTR %)",
    color = "Ad Format"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

# Practical Recommendations

1. **Update Product Catalog Images Every 2 Weeks:**
   The `Product_Catalog_Retargeting` campaign loses clicks quickly once a user sees the ad more than **1.5 times**. Updating product frames or changing background colors every 14 days will keep click rates high.

2. **Use Video Views to Feed Re-targeting Audiences:**
   The `Brand_Video_Views` campaign has a lower direct sales return, but it reaches people at the lowest cost. Building a custom audience of users who watched at least 50% of these videos and showing them special offers will lower customer acquisition costs.

3. **Reallocate 15% of the Budget to Re-targeting:**
   Move around 15% of the budget from broad discovery carousels to the product catalog re-targeting campaigns, which consistently generate over **3.0x ROAS**.
