---
title: "FDS Final Project: Report #4"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(readxl)
library(knitr)
library(ggplot2)
library(infer)
library(janitor)
```


# Introduction 


## The Data from the spreadsheet


```{r}
athletes <- read_excel("olympics.xlsx", sheet = "athletes")
country <- read_excel("olympics.xlsx", sheet = "country")
games <- read_excel("olympics.xlsx", sheet = "games")
medals <- read_excel("olympics.xlsx", sheet = "medals")

glimpse(athletes)
glimpse(country)
glimpse(games)
glimpse(medals)

summary(athletes)
summary(country)
```

# Data Analysis


## Part 1. Athletes info


```{r}
multi_country_athletes <- country %>%
  group_by(athlete_id) %>%
  summarise(n_countries = n_distinct(NOC)) %>%
  filter(n_countries > 1) %>%
  nrow()
```


_In history of the Olympic Games **`r multi_country_athletes`** athletes have competed for more than one country_



# Who are the ten athletes that took part in most games?


## Part 2.


```{r}
top_games <- country %>%
  group_by(athlete_id) %>%
  summarise(n_games = n_distinct(Games)) %>%
  arrange(desc(n_games)) %>%
  head(10) %>%
  left_join(athletes, by = c("athlete_id" = "ID"))

top_games %>%
  select(Name, n_games) %>% 
  kable(caption = "Top 10 Athletes by Number of Olympic Games Participated")
```


# Who are the 10 athletes that competed in the most events?


## Part 3.


```{r}
top_events <- medals %>%
  group_by(athlete_id) %>%
  summarise(n_events = n()) %>%
  arrange(desc(n_events)) %>%
  head(10) %>%
  left_join(athletes, by = c("athlete_id" = "ID"))

top_events %>%
  select(Name, n_events) %>% 
  kable(caption = "Top 10 Athletes by Number of Events Competed")

# Checking if one athlete competed more than once
multi_event_check <- medals %>%
  group_by(athlete_id, Games) %>%
  summarise(n_events_game = n(), .groups = "drop") %>%
  filter(n_events_game > 1) %>%
  nrow()
```


**And Yes**, _many athetes have competed more than once (in total **`r multi_event_check`**)_


# Countries info


## Part 4. Table of 15 countries with medals


```{r}
medals_full <- medals %>%
  left_join(country, by = c("athlete_id", "Games")) %>%
  left_join(games, by = "Games")

top_15_nocs <- medals_full %>%
  filter(!is.na(Medal)) %>%
  count(NOC, sort = TRUE) %>%
  head(15) 

medal_matrix <- medals_full %>%
  filter(!is.na(Medal)) %>%
  inner_join(top_15_nocs, by = "NOC") %>%
  count(NOC, Year, name = "Total_Medals") %>%
  pivot_wider(names_from = Year, values_from = Total_Medals, values_fill = 0) %>%
  # column with total medals
  mutate(num_cols = select(., where(is.numeric)),
         Total = rowSums(num_cols)) %>% 
  select(-num_cols)
  

medal_matrix %>% 
  kable(caption = "Medal Count Matrix by Country and Year (Top 15 Countries)")

medal_success <- medals_full %>%
  inner_join(top_15_nocs, by = "NOC") %>%
  mutate(won_medal = if_else(!is.na(Medal), "Won", "No_Medal"))

success_rate <- medal_success %>%
  group_by(NOC) %>%
  summarise(total_competed = n(),
    total_won = sum(!is.na(Medal)),
    success_pct = (total_won / total_competed) * 100)

success_rate %>% 
  kable(caption = "Medal Succes Rate by country (top 15 countries")
```


# Is there a relationship between countries?

## Part 5.

```{r}

ggplot(success_rate, aes(x = reorder(NOC, success_pct), y = success_pct)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Medal Success Percentage by Country (Top 15 Countries)",
    x = "Country (NOC)",
    y = "Percentage of Medals Won (%)") +
  theme_minimal()
```

**The chart shows that medal chances change by country**

_URS and GDR show the highest success rates, reaching around 40%, meaning they earned medals in a large share of the events they entered_.

_USA is next at about 30%, followed by RUS and GER in the low 20s_.

_FRA and CAN are in the lower range, with success rates around 13–15%_.


## The Chi Square test

```{r}
crosstab_table <- table(medal_success$NOC, medal_success$won_medal)
chi_result <- chisq.test(crosstab_table)
chi_result
```

**This result confirms what the bar chart already suggested:** _Countries like URS and GDR reach rates near 40%, while others stay around 13–15%. The data shows that success is not only about how many events a country enters, but that medal probability varies clearly from one country to another_


# Athletes and Sports

## Part 6. 

```{r}
sport_data <- medals_full %>%
  left_join(athletes, by = c("athlete_id" = "ID")) %>%
  group_by(Sport) %>%
  summarise(height = mean(Height, na.rm = TRUE),
    weight = mean(Weight, na.rm = TRUE)) %>%
  mutate(bmi = weight / ((height / 100)^2)) %>%
  filter(!is.na(height) & !is.na(weight))

extremes <- sport_data %>%
  mutate(label = case_when(
      height == max(height, na.rm = TRUE) ~ paste0(Sport, ": Max Height"),
      height == min(height, na.rm = TRUE) ~ paste0(Sport, ": Min Height"),
      weight == max(weight, na.rm = TRUE) ~ paste0(Sport, ": Max Weight"),
      weight == min(weight, na.rm = TRUE) ~ paste0(Sport, ": Min Weight"),
      bmi == max(bmi, na.rm = TRUE) ~ paste0(Sport, ": Max BMI"),
      bmi == min(bmi, na.rm = TRUE) ~ paste0(Sport, ": Min BMI"),
      TRUE ~ NA_character_)) %>%
  filter(!is.na(label))

# Plot
sport_data %>%
  ggplot(aes(x = weight, y = height)) +
  geom_point(color = "gray", alpha = 0.7) +
  geom_point(data = extremes, color = "red", size = 3) +
  geom_text(data = extremes, aes(label = label), vjust = -1, size = 3) + 
  labs(title = "Average Height vs Weight by Sport",
    x = "Weight (kg)",
    y = "Height (cm)") +
  theme_minimal()
```


# Plot showing the number of medals


## Part 7.


```{r}

medals_by_year <- medals_full %>%
  filter(!is.na(Medal)) %>%
  # eliminate the spike of medals in 1956
  distinct(athlete_id, Games, Event, Medal, .keep_all = TRUE) %>% 
  group_by(Year, Season, Medal) %>%
  summarise(total_medals = n(), .groups = "drop")


medals_by_year %>%
  ggplot(aes(x = Year, y = total_medals, color = Medal)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  facet_wrap(~ Season, scales = "free_y") +
  scale_color_manual(values = c("Gold" = "gold", "Silver" = "gray", "Bronze" = "orange")) +
  labs(title = "Number of Medals Awarded by Year",
    x = "Year",
    y = "Total Medals",
    color = "Medal Type") +
  theme_minimal()
```

**Summer Games have many more medals and show peaks, while Winter Games grow slowly and steadily**



