---
title: "FDS Final Project: Report #3"
output:
  html_document:
    df_print: paged
---

# Final Project # 3.

## Configuration and Set UP

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
library(broom)

# API Configuration
api_key <- "620d2f84e10959ab395497ae282f69cb" # Personal API key
base_url <- "https://api.themoviedb.org/3"
```


# Specific Queries

## Part 1. Extraction of Data

```{r}
# Highest-Grossing Dramas from 2010
url_dramas <- paste0(base_url, "/discover/movie?api_key=", api_key, "&with_genres=18&primary_release_year=2010&sort_by=revenue.desc")
tibble_dramas <- content(GET(url_dramas), as = "parsed", simplifyVector = TRUE)$results %>% as_tibble()
write_csv(tibble_dramas, "dramas_2010.csv")

# Will Ferrell and Liam Neeson Collaboration
url_fn <- paste0(base_url, "/discover/movie?api_key=", api_key, "&with_cast=23659,3896")
tibble_fn <- content(GET(url_fn), as = "parsed", simplifyVector = TRUE)$results %>% as_tibble()
write_csv(tibble_fn, "ferrell_neeson.csv")

# Kids' Movies Featuring Tom Cruise
url_cruise <- paste0(base_url, "/discover/movie?api_key=", api_key, "&with_cast=500&certification_country=US&certification.lte=G")
tibble_cruise <- content(GET(url_cruise), as = "parsed", simplifyVector = TRUE)$results %>% as_tibble()
write_csv(tibble_cruise, "tom_cruise_kids_movies.csv")
```

### Dramas from 2010: Films like INCEPTION They top the list with the highest earnings of that year in the drama genre.

### Will Ferrell and Liam Neeson: They share credits in animated productions such as The Lego Movie, where they appeared as voice actors, and these entries can be verified in the tibble.

### As for Tom Cruise: The tibble shows not exactly for kids but for all public


# Pixar ID Extraction

## Part 2. Pixar ID company

```{r}
url_search_pixar <- paste0(base_url, "/search/company?api_key=", api_key, "&query=Pixar")

pixar_id <- content(GET(url_search_pixar), as = "parsed", simplifyVector = TRUE)$results %>% 
  filter(grepl("Pixar", name, ignore.case = TRUE)) %>% 
  pull(id) %>% 
  first()
```

### This is the Pixar ID `r pixar_id`


# All Pixar Movies

## Part 3. Extraction Code 

```{r}
# Define a function to fetch a specific page of Pixar movies
get_pixar <- function(p) {
  url_pixar <- paste0(base_url, "/discover/movie?api_key=", api_key, "&with_companies=", pixar_id, "&sort_by=revenue.desc&page=", p)
  content(GET(url_pixar), as = "parsed", simplifyVector = TRUE)$results
}

# Fetch multiple pages (e.g., pages 1 to 5) and bind them into a single tibble
tibble_pixar_all <- map(1:5, get_pixar) %>% 
  bind_rows() %>% 
  as_tibble()

# View the first few rows of the final tibble
head(tibble_pixar_all)
tibble_pixar_all
```


### Output Summary: The tibble condenses the dataset into one film per row with 15 key metadata columns, and the preview shows top‑grossing titles like Inside Out 2 and Incredibles 2 leading the sorted list.


# Comparative Analysis Pixar vs. Disney from 2006

## Part 4.

### 4.1 Finding Wald Disney ID

```{r}
url_search_disney <- paste0(base_url, "/search/company?api_key=", api_key, "&query=Walt%20Disney%20Pictures")

# Extract the ID safely
disney_id <- content(GET(url_search_disney), as = "parsed", simplifyVector = TRUE)$results %>% 
  filter(grepl("Walt Disney Pictures", name, ignore.case = TRUE)) %>% 
  pull(id) %>% 
  first()
```

### The Wald Disney ID is `r disney_id`

### 4.2 Fetching Movies from 2006 Onwards for Both Studios

```{r}
get_studio_movies <- function(company_id, max_pages = 5) {
  map(1:max_pages, function(p) {
    url <- paste0(base_url, "/discover/movie?api_key=", api_key, 
                  "&with_companies=", company_id, 
                  "&primary_release_date.gte=2006-01-01",
                  "&sort_by=revenue.desc&page=", p)
    res <- content(GET(url), as = "parsed", simplifyVector = TRUE)
    if (!is.null(res$results) && length(res$results) > 0) {
      res$results %>% mutate(studio = ifelse(company_id == 3, "Pixar", "Disney"))
    } else {
      NULL
    }
  }) %>% bind_rows()
}

# Retrieve data for both studios
pixar_2006 <- get_studio_movies(pixar_id, max_pages = 3)
disney_2006 <- get_studio_movies(disney_id, max_pages = 5)

# Combine and clean data (using 'release_date' in stead of 'primary_release_date')
combined_movies <- bind_rows(pixar_2006, disney_2006) %>% 
  filter(vote_count >= 50) %>% 
  select(title, vote_average, vote_count, studio, release_date)
```


### 4.3 Boxplot Comparition

```{r}
ggplot(combined_movies, aes(x = studio, y = vote_average, fill = studio)) +
  geom_boxplot(alpha = 0.7, outlier.colour = "red", outlier.shape = 8) +
  labs(
    title = "Vote Average Comparison: Pixar vs. Disney (2006+)",
    x = "Studio",
    y = "Vote Average (TMDB)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Pixar films show stronger and more consistent audience and critic approval after the acquisition, reinforcing that Disney’s purchase was a solid strategic move for sustaining higher level animation quality.

### 4.4 T-Test

```{r}
t_test_result <- t.test(vote_average ~ studio, data = combined_movies)
t_test_result
```
### Statistical Hypothesis: The Welch test shows a clear and significant difference between studios: Pixar’s mean score is 7.07 versus Disney’s 6.55, and the very small p‑value (0.0001021) confirms the gap isn’t due to chance.

### Final Analysis: Disney’s 2006 acquisition of Pixar proved to be a smart strategic move. Even after years of collaboration, fully integrating Pixar gave Disney a reliable engine of creative consistency. The data shows that Pixar films kept higher and more stable audience and critic ratings post‑acquisition, reinforcing that the purchase was a strong long‑term investment in animation quality.





