---
title: "FDS Final Project: Report #3"
output:
  html_document:
    df_print: paged
---

# Final Project # 3

## Configuration and Set UP


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(httr)
library(jsonlite)
library(broom)
library(knitr)

# API Configuration
api_key <- "620d2f84e10959ab395497ae282f69cb"
base_url <- "https://api.themoviedb.org/3"
```


# Specific Queries

## Part 1. Extraction of Data

```{r}
# Highest dramas from 2010
res_dramas <- GET(paste0(base_url, "/discover/movie"),
  query = list(api_key = api_key,
    with_genres = "18",
    primary_release_year = "2010",
    sort_by = "revenue.desc"))

tibble_dramas <- res_dramas %>%
  content(as = "parsed", simplifyVector = TRUE) %>%
  pluck("results") %>%
  as_tibble()

# Will and Liam 
res_fn <- GET( paste0(base_url, "/discover/movie"),
  query = list(api_key = api_key,
    with_cast = "23659,3896"))

tibble_fn <- res_fn %>%
  content(as = "parsed", simplifyVector = TRUE) %>%
  pluck("results") %>%
  as_tibble()

# Kids Movies Featuring Tom Cruise
res_cruise <- GET(paste0(base_url, "/discover/movie"),
  query = list(api_key = api_key,
    with_cast = "500",
    certification_country = "US",
    "certification.lte" = "G"))

tibble_cruise <- res_cruise %>%
  content(as = "parsed", simplifyVector = TRUE) %>%
  pluck("results") %>%
  as_tibble()
```

## Preparing for the inline code

```{r}
tibble_dramas %>% 
  select(id, title, vote_average, vote_count, release_date) %>% 
  head(5) %>% 
  kable(caption = "Top 5 Dramas from 2010")

tibble_fn %>% 
  select(id, title, vote_average, vote_count, release_date) %>% 
  head(5) %>% 
  kable(caption = "Movies featuring Will Ferrell and Liam Neeson")

tibble_cruise %>% 
  select(id, title, vote_average, vote_count, release_date) %>% 
  head(5) %>% 
  kable(caption = "Tom Cruise Family Movies")
```

_Top drama films of 2010_ **`r tibble_dramas %>% slice(1) %>% pull(title)`**.

_Collaborated Production like **`r tibble_fn %>%  slice(1) %>%  pull(title)`**, between Will Ferrel and Liam Neeson_.

_Featuring Tom Cruise such as **`r tibble_cruise %>% slice(1) %>%  pull(title)`** not only for kids but categorized for all audience_.



# Pixar ID Extraction

## Part 2. Pixar ID company

```{r}
url_search_pixar <- paste0(base_url, "/search/company?api_key=", api_key, "&query=Pixar")

pixar_id <- GET(url_search_pixar) %>% 
  content(as = "parsed", simplifyVector = TRUE) %>% 
  pluck("results") %>% 
  filter(grepl("Pixar", name, ignore.case = TRUE)) %>% 
  pull(id) %>% 
  first()
```

_This is the Pixar ID_ **`r pixar_id`**


# All Pixar Movies

## Part 3. Extraction Code 

```{r}

# Getting total pages 
url_pixar_page1 <- paste0(base_url, "/discover/movie?api_key=", api_key, "&with_companies=", pixar_id, "&sort_by=revenue.desc&page=1")
res_pixar_p1 <- GET(url_pixar_page1) %>%
  content(as = "parsed", simplifyVector = TRUE)
total_pages_pixar <- res_pixar_p1 %>%
  pluck("total_pages")

get_pixar <- function(p) {
  url <- paste0(base_url, "/discover/movie?api_key=", api_key, "&with_companies=", pixar_id, "&sort_by=revenue.desc&page=", p)
  GET(url) %>% content(as = "parsed", simplifyVector = TRUE) %>% pluck("results")
}

# Binded tibble
tibble_pixar_all <- map(1:total_pages_pixar, get_pixar) %>% 
  bind_rows() %>% 
  as_tibble()

# Preview top results
tibble_pixar_all %>% 
  select(id, title, vote_average, vote_count, release_date) %>% 
  head(5) %>% 
  kable(caption = "Top Pixar Movies")
```

**Output Summary:** _The dataset contains **`r nrow(tibble_pixar_all)`** Pixar films featuring top titles of the sorted list_


# Comparative Analysis Pixar vs. Disney from 2006

## Part 4.

### 4.1 Finding Wald Disney ID

```{r}
url_search_disney <- paste0(base_url, "/search/company?api_key=", api_key, "&query=Walt%20Disney%20Pictures")

disney_id <- GET(url_search_disney) %>% 
  content(as = "parsed", simplifyVector = TRUE) %>% 
  pluck("results") %>% 
  filter(grepl("Walt Disney Pictures", name, ignore.case = TRUE)) %>% 
  pull(id) %>% 
  first()
```

_The Wald Disney ID is_ **`r disney_id`**


### 4.2 Fetching Movies from 2006 Onwards for Both Studios

```{r}

get_studio_movies <- function(company_id) {
  url_p1 <- paste0(base_url, "/discover/movie?api_key=", api_key, 
                   "&with_companies=", company_id, 
                   "&primary_release_date.gte=2006-01-01",
                   "&sort_by=revenue.desc&page=1")
  res_p1 <- GET(url_p1) %>% content(as = "parsed", simplifyVector = TRUE)
  tot_pages <- res_p1 %>%
    pluck("total_pages")
  if (is.null(tot_pages) || tot_pages == 0) tot_pages <- 1
  
  # Retrieving all data 
  map(1:tot_pages, function(p) {
    url <- paste0(base_url, "/discover/movie?api_key=", api_key, 
                    "&with_companies=", company_id, 
                    "&primary_release_date.gte=2006-01-01",
                    "&sort_by=revenue.desc&page=", p)
    res <- GET(url) %>%
      content(as = "parsed", simplifyVector = TRUE)
    results <- res %>%
      pluck("results")
    if (!is.null(results) && length(results) > 0) {results %>%
        mutate(studio = ifelse(company_id == pixar_id, "Pixar", "Disney"))} else { NULL}
  }) %>%
    bind_rows()}

pixar_2006 <- get_studio_movies(pixar_id)
disney_2006 <- get_studio_movies(disney_id)

# Excluding Pixar from Disney movies
disney_clean <- disney_2006 %>%
  filter(!id %in% pixar_2006$id)
pixar_clean <- pixar_2006

# combined data set and cleaned
combined_movies <- bind_rows(pixar_clean, disney_clean) %>% 
  filter(vote_count >= 50) %>% 
  select(id, title, vote_average, vote_count, studio, release_date)
```


### 4.3 Boxplot Comparition

```{r}
combined_movies %>% 
  ggplot(aes(x = studio, y = vote_average, fill = studio)) +
  geom_boxplot(alpha = 0.7, outlier.colour = "orange", outlier.shape = 8) +
  labs(title = "Vote Average Comparison: Pixar vs. Disney (2006+)",
    x = "Studio",
    y = "Vote Average (TMDB)") +
  theme_minimal() +
  theme(legend.position = "none")
```

**Pixar films show stronger and more consistent audience and critic approval after the acquisition, reinforcing that Disney’s purchase was a solid strategic move for sustaining higher level animation quality.**

### 4.4 T-Test

```{r}
library(infer)

# T-Test with {infer} package
t_test_result <- combined_movies %>% 
  t_test(response = vote_average, explanatory = studio, order = c("Pixar", "Disney"))

t_test_result
```
**Statistical Hypothesis:** _The t-test confirms a statistically significant difference between the studios mean scores, with Disney outperforming Pixar, The very small p-value indicates that this gap is reliable and not due to chance, supporting the long term creative and strategic success of Disney's acquisition._

### Final Analysis: Disney’s 2006 acquisition of Pixar proved to be a smart strategic move. Even after years of collaboration, fully integrating Pixar gave Disney a reliable engine of creative consistency. The data shows that Pixar films kept higher and more stable audience and critic ratings post‑acquisition, reinforcing that the purchase was a strong long‑term investment in animation quality.





